---
title: "Databases w/ R"
output: html_notebook
---

```{r}
library(DBI)
library(odbc)

con <- dbConnect(odbc(), 
                      Driver = "SQL Server", 
                      Server = "localhost\\SQLEXPRESS", 
                      Database = "master", 
                      Trusted_Connection = "True")
```


```{r}
dbGetQuery(con, "Select Top 10 * From flights")
```

```{sql, connection = con}
Select top 10 * from "flights"
```

## dplyr & dbplyr

Install `dbplyr`, along with `dplyr`, to use interact with databases

```{r, eval = FALSE}
devtools::install_github("tidyverse/dbplyr")
```

```{r}
library(dplyr)
library(dbplyr)

```


```{r}
flights <- tbl(con, "flights")
```

```{r}
object.size(flights)
```

```{r}
flights

```


```{r}
flights %>%
  tally()
```

```{r}
flights %>%
  group_by(month) %>%
  tally() 
```

### show_query()

```{r}
flights %>%
  group_by(name) %>%
  tally() %>%
  show_query()
```


### join tables

```{r}
airports <- tbl(con, "airports")

```

```{r}
new_flights <- flights %>%
  left_join(airports, by = c("dest" = "faa"))

new_flights
```

### collect()

```{r}
by_location <- new_flights %>%
  group_by(name, lon, lat) %>%
  tally() %>%
  collect()

by_location 
```



### plotting

```{r, fig.width = 10, fig.height = 5}
library(ggplot2) 

by_location %>%
  ggplot() +
  geom_point(aes(x = lon, y= lat, size = n, color = n), alpha = 0.5)

```

### Direct insights

```{r}
new_flights %>%
  group_by(name) %>%
  tally() %>%
  arrange(desc(n)) %>%  # desc() inside arrange() get the top counts
  head(10)

```

### pull() command

```{r}
top_names <- new_flights %>%
  group_by(name) %>%
  tally() %>%
  arrange(desc(n)) %>%
  head(10) %>% 
  collect() %>%
  pull("name")

top_names
```

### Take advantage of dplyr's laziness

```{r}
top_flights <- new_flights %>%
  filter(name %in% top_names)

```

Use `View()` to preview some of the data in the familiar data viewer

```{r}
top_flights %>%
  head(1000) %>%
  collect() %>%
  View("top_flights")
```

Easily transition from data manipulation to plotting, without needing to import the data

```{r}
top_flights %>%
  group_by(Month) %>%
  tally() %>%
  collect() %>%
  ggplot() +
    geom_area(aes(x = Month , y = n), fill = "blue", alpha = 0.5)
  
```


Thanks to the `tidyverse` approach, an entirly new plot can be built by adding one single line, and the data is still summarized inside the database

```{r}
top_flights %>%
  filter(distance > 2000) %>%  # Added 
  group_by(Month) %>%
  tally() %>%
  collect() %>%
  ggplot() +
    geom_area(aes(x = Month , y = n), fill = "blue", alpha = 0.5)
  
```

### Histograms

```{r}
library(dbplot)

flights %>%
  dbplot_histogram(arr_delay, binwidth = 20)

```
## Temporary tables

```{r}
temp_flights <- new_flights %>%
  group_by(name) %>%
  tally() 
```

```{r}
temp_flights <- copy_to(con, temp_flights, "temp_flights")
```


```{r}
temp_flights %>%
  arrange(desc(n))
```

```{r}
dbDisconnect(con)
```




