---
title: "Austin - 311 Calls"
resource_files:
- .httr-oauth
output: html_notebook
---

```{r setup, include = FALSE}
library(DBI)
library(dbplot)
library(bigrquery)
library(dplyr, warn.conflicts = FALSE)
library(dbplyr, warn.conflicts = FALSE)
library(modeldb)
library(leaflet)
```

## Background

City of Austin 3-1-1 service for Police non-emergency calls.  This helped to reduce the number of calls to 9-1-1 so they could concentrate on those calls where either life and /or property were in danger.
 
 - Link: http://www.austintexas.gov/department/about-3-1-1

## BigQuery 

Google BigQuery provides access to the 3-1-1 data as part of their Public Data project: 	
`bigquery-public-data:austin_311.311_service_requests`

```{r}
library(DBI)
library(bigrquery)

con <- dbConnect(
  bigquery(),
  project = "bigquery-public-data",
  dataset = "austin_311",
  billing = "rstudio-bigquery-event",
  use_legacy_sql = FALSE
)
```

## dplyr with BigQuery

Instead of importing the entire data set, we just create a pointer to the table inside BigQuery:
```{r}
service <- tbl(con, "311_service_requests")
```

Standard `dplyr` commands work inside BigQuery:
```{r}
glimpse(service)
```

```{r}
tally(service)
```

Which are the most common complaints?
```{r}
service %>%
  count(complaint_description) %>%
  top_n(10)
```

Some data cleaning...
```{r}
service <- service %>%
  filter(latitude > 10)
```

## Visualizations using dbplot, ggplot2 and leaflet

Easily view the location of the service calls without importing all of the data.
```{r}
library(dbplot)

complaint <- "Loose Dog"

service %>%
  filter(complaint_description == complaint) %>%
  dbplot_raster(longitude, latitude, resolution = 50)
  
```

Use the new `db_computer_raster2()` function to obtain the pre-calculated squares, with their limits included.
```{r}
calls <- service %>%
  filter(complaint_description == complaint) %>% 
  db_compute_raster2(longitude, latitude, resolution = 30) %>%
  mutate(of_max = `n()` / max(`n()`))

calls
```

Overlay the squares over a map of Austin using `leaflet`
```{r, fig.width = 8}
library(leaflet)

leaflet() %>% 
  addTiles() %>%
  addRectangles(
    calls$longitude, 
    calls$latitude, 
    calls$longitude_2, 
    calls$latitude_2
  )
```

Make it fancy... Highlight squares based on number of comparative number of calls, and add a popup
```{r, fig.width = 8}

fancy <- leaflet() %>% 
  addTiles() %>%
  addRectangles(
    calls$longitude, 
    calls$latitude, 
    calls$longitude_2, 
    calls$latitude_2,
    fillOpacity = calls$of_max,
    fillColor = "red",
    stroke = FALSE,
    popup = paste0("<p>Calls: ",  calls$`n()`,"</p>")
  )

fancy
```

Where's the Fairmont?
```{r, fig.width = 8}
fancy  %>%
  addMarkers(lng = -97.7404001, lat = 30.2620562)
```

## Kmeans with modeldb

Calculate k-means inside the database using `simple_kmeans_db()`
```{r}
service_kmeans <-  service %>%
  filter(complaint_description == complaint) %>%
  simple_kmeans_db(latitude, longitude)

head(service_kmeans, 10)
```

Preview the final SQL statement that the routine settled on
```{r}
show_query(service_kmeans)
```

Create a local table with the corners of each center
```{r}
cr <- service_kmeans %>%
  group_by(k_longitude, k_latitude) %>%
  summarise(
    calls = n(),
    lat1 = min(latitude, na.rm = TRUE),
    lat2 = max(latitude, na.rm = TRUE),
    lon1 = min(longitude, na.rm = TRUE),
    lon2 = max(longitude, na.rm = TRUE),
  ) %>%
  collect()
  
cr
```

Plot against a map
```{r, fig.width = 8}
leaflet() %>% 
  addTiles() %>%
  addRectangles(cr$lon1, cr$lat1, cr$lon2, cr$lat2) 
```

Overlay the raster boxes
```{r, fig.width = 8}
leaflet() %>% 
  addTiles() %>%
  addRectangles(cr$lon1, cr$lat1, cr$lon2, cr$lat2) %>%
  addRectangles(
    calls$longitude, 
    calls$latitude, 
    calls$longitude_2, 
    calls$latitude_2,
    fillColor = "red",
    stroke = FALSE,
    popup = paste0("<p>Calls: ",  calls$`n()`,"</p>")
  )  
  
```

## Bringing it all together 

Use `purrr` to run `db_compute_raster()` for each center in the kmeans results
```{r}
library(purrr)

cents <- c("center_1", "center_2", "center_3")

by_centers <- cents %>%
  map_df(~{
    bc <- service_kmeans %>% 
      filter(k_center == .x) %>%
      db_compute_raster2(longitude, latitude, resolution =  20) %>%
      mutate(
        of_max = `n()` / max(`n()`),
        center = .x
        )
    })
```

Change the color of each center
```{r}
bc <- by_centers %>%
  mutate(color = case_when(
    center == "center_1" ~ "red",
    center == "center_2" ~ "blue",
    center == "center_3" ~ "green",
    TRUE ~ "white"
  ))

bc
```

Map the two layers together
```{r, fig.width = 8}
leaflet() %>% 
  addTiles()  %>%
  addRectangles(cr$lon1, cr$lat1, cr$lon2, cr$lat2) %>%
  addRectangles(
    bc$longitude, 
    bc$latitude, 
    bc$longitude_2, 
    bc$latitude_2,
    fillColor = bc$color,
    fillOpacity = 0.5,
    stroke = FALSE,
    popup = paste0("<p>Calls: ",  bc$`n()`,"</p>")
  )
```

```{r}
dbDisconnect(con)
```

